Quantum Interval-Valued Probability: Measurement and Probability

### 2017-11-30 prop_iop.tex
prop_iop.tex \section{Expectation Values for IVPMs} Updating...

\begin{definition}[Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}]\label{def:Choquet}~
\begin{itemize}
\item Consider a classical IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
\item We still decompose a random variable~$X$ into step functions, but
we separate the positive and negative terms and write like $\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}+\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}$,
i.e., 
\begin{equation}
x_{N^{-}}^{-}<\ldots<x_{1}^{-}<0\le x_{1}^{+}<\ldots<x_{N^{+}}^{+}
\end{equation}
 and $\left\{ E_{i}^{-}\right\} _{i=1}^{N^{-}}$ and $\left\{ E_{i}^{+}\right\} _{i=1}^{N^{+}}$
are all disjoint.
\item Let $E_{i}^{s\prime}=\bigcup_{j\ge i}E_{j}^{s}$ and $\Delta\mu^{*}\left(E_{i}^{s\prime}\right)=\mu^{*}\left(E_{i}^{s\prime}\right)-\mu^{*}\left(E_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}\eqalign{ 
\int X\rmd\bar{\mu} = & \left[\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mul\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mur\left(E_{i}^{-\prime}\right),\right.\\  
& \left.\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mur\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mul\left(E_{i}^{-\prime}\right)\right]\,.
}\label{eq:interval-expectation}\end{equation}
\end{definition}

@@

Before discussing QIVPMs, we first review various notations brought
forward from conventional quantum theory. Assume that a physical system
is represented by a density matrix~$\rho$, and that we wish to measure
a physical quantity~${\cal O}$ of this system, with corresponding
observable~$\mathbf{O}$, whose spectral decomposition is given by~\cite{Griffiths1994,FraleighBeauregard1995,544199,FriedbergInselSpence2002,Jaeger2007}
\begin{equation}
\mathbf{O}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}P_{i}^{s}\,,\label{eq:spectrum-decomposition-distinct}
\end{equation}
where the positive and negative eigenvalues are ordered from the smallest
to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}<\ldots<\lambda_{1}^{-}<0\le\lambda_{1}^{+}<\ldots<\lambda_{N^{+}}^{+}\,,
\end{equation}
and $P_{i}$ is the projector onto the eigenspace of each distinct
eigenvalue~$\lambda_{i}$. Compare to equation~(\ref{eq:spectrum-decomposition-distinct}),
the spectral decomposition of $\mathbf{O}$ may be represented in
term of outer product as $\mathbf{O}=\sum_{k=1}^{d}\lambda_{k}\proj{k}$
which would be used in lemma~\ref{lemma:observable2random-variable}
later. Disregard the representation of the spectral decomposition,
it can be used to defined the expectation value of the observable~$\mathbf{O}$
as~\cite{544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}=\Tr\left(\rho\mathbf{O}\right)=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mu_{\rho}^{\mathrm{B}}\left(P_{i}^{s}\right)\ .\label{eq:quantum-expectation}
\end{equation}

Consider a QIVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$. Similar to the classical Choquet integral
in definition~\ref{def:Choquet}, we can let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$
and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $,
and define the expectation value of $\mathbf{O}$ with respect to
$\bar{\mu}$ as
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\left[\Delta\mul\left(P_{i}^{s\prime}\right),\Delta\mur\left(P_{i}^{s\prime}\right)\right]\,.\label{eq:quantum-interval-expectation}
\end{equation}

\subsection{Between Expectation Values for Classical and Quantum IVPMs\label{subsec:Between-Expectation-Values-Classical-Quantum}}

Since the similarity between classical and quantum expectation values,
we would prove some lemmas\footnote{\yutsung{functor? Galois connection?}}
which can help us move the classical results to quantum.

>>>>

\begin{definition}[Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}]\label{def:Choquet}~
\begin{itemize}
\item Consider a classical IVPM~$\bar{\mu}$ such that $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$ are the left-end and
the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\mul\left(E\right),\mur\left(E\right)\right]$.
\item We still decompose a random variable~$X$ into step functions, but
we separate the positive and negative terms and write 
\begin{equation}
X=\sum_{i=1}^{N^{-}}x_{i}^{-}\mathbf{1}_{E_{i}^{-}}+\sum_{i=1}^{N^{+}}x_{i}^{+}\mathbf{1}_{E_{i}^{+}}\,,\label{eq:classical-decomposition}
\end{equation}
where 
\begin{equation}
x_{N^{-}}^{-}<\ldots<x_{1}^{-}<0\le x_{1}^{+}<\ldots<x_{N^{+}}^{+}
\end{equation}
and $\left\{ E_{i}^{-}\right\} _{i=1}^{N^{-}}$ and $\left\{ E_{i}^{+}\right\} _{i=1}^{N^{+}}$
are all disjoint.
\end{itemize}
Then, the Choquet integral of $X$ with respect to $\bar{\mu}$ is
\begin{equation}\eqalign{ 
\int X\rmd\bar{\mu} = & \left[\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mul\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mur\left(E_{i}^{-\prime}\right),\right.\\  
& \left.\sum_{i=1}^{N^{+}}x_{i}^{+}\Delta\mur\left(E_{i}^{+\prime}\right)+\sum_{i=1}^{N^{-}}x_{i}^{-}\Delta\mul\left(E_{i}^{-\prime}\right)\right]\,.
}\label{eq:interval-expectation}\end{equation}
\end{definition}

@@

Before discussing QIVPMs, we first review various notations brought
forward from the conventional quantum theory. Assume that we wish
to measure a physical quantity~${\cal O}$ of this system, with corresponding
observable~$\mathbf{O}$. Its spectral decomposition can either be
represented in term of outer product or projector. Since we want to
define the expectation value of~$\mathbf{O}$ with respect to a QIVPM
as the classical Choquet integral in definition~\ref{def:Choquet},
we are going to order the eigenvalues of~$\mathbf{O}$ as we ordered
the step functions for classical random variables, and to represent
the spectral decomposition of~$\mathbf{O}$ by~\cite{Griffiths1994,FraleighBeauregard1995,544199,FriedbergInselSpence2002,Jaeger2007}
\begin{equation}
\mathbf{O}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}P_{i}^{s}\,,\label{eq:spectrum-decomposition-distinct}
\end{equation}
where the positive and negative eigenvalues are ordered from the smallest
to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}<\ldots<\lambda_{1}^{-}<0\le\lambda_{1}^{+}<\ldots<\lambda_{N^{+}}^{+}\,,
\end{equation}
and $P_{i}^{s}$ is the projector onto the eigenspace of each distinct
eigenvalue~$\lambda_{i}^{s}$.

For the conventional quantum theory, a physical system can be represented
by a density matrix~$\rho$, the expectation value of the observable~$\mathbf{O}$
is defined as~\cite{544199,Jaeger2007}
\begin{equation}
\expval{\mathbf{O}}_{\mu_{\rho}^{\mathrm{B}}}=\Tr\left(\rho\mathbf{O}\right)=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\mu_{\rho}^{\mathrm{B}}\left(P_{i}^{s}\right)\ .\label{eq:quantum-expectation}
\end{equation}
Equation~(\ref{eq:quantum-expectation}) can be combined with definition~\ref{def:Choquet}
to define the expectation value of $\mathbf{O}$ with respect to a
QIVPM~$\bar{\mu}$. Similar to definition~\ref{def:Choquet}, we
first separate the left-end and the right-end of $\bar{\mu}$ as $\mul:\events\rightarrow\left[0,1\right]$
and $\mur:\events\rightarrow\left[0,1\right]$. Then, let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$
and $\Delta\mu^{*}\left(P_{i}^{s\prime}\right)=\mu^{*}\left(P_{i}^{s\prime}\right)-\mu^{*}\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $\mu^{*}\in\left\{ \mul,\mur\right\} $.
Finally, the expectation value of $\mathbf{O}$ with respect to a
QIVPM~$\bar{\mu}$ is defined to be
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\sum_{s\in\left\{ +,-\right\} }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}\left[\Delta\mul\left(P_{i}^{s\prime}\right),\Delta\mur\left(P_{i}^{s\prime}\right)\right]\,.\label{eq:quantum-interval-expectation}
\end{equation}

Notice that the similarity between classical Choquet integral and
quantum expectation values as in Table~\ref{tab:compare-CIVPM-QRVPM}.
\begin{table}
\caption{\label{tab:compare-CIVPM-QRVPM}Compare the similarity between the
classical and quantum}
\centering{}%
\begin{tabular}{cc}
\toprule 
\addlinespace
Classical & Quantum\tabularnewline\addlinespace
\midrule
\midrule 
\addlinespace
CIVPM~$\bar{\mu}$ & QIVPM~$\bar{\mu}$\tabularnewline\addlinespace
\midrule 
\addlinespace
Random variable~$X$ & Observable~$\mathbf{O}$\tabularnewline\addlinespace
\midrule 
\addlinespace
Classical step function decomposition, equation~(\ref{eq:classical-decomposition}) & Spectral decomposition, equation~(\ref{eq:spectrum-decomposition-distinct})\tabularnewline\addlinespace
\midrule 
\addlinespace
Distinct value~$x_{i}^{s}$ & Distinct eigenvalue~$\lambda_{i}^{s}$\tabularnewline\addlinespace
\midrule 
\addlinespace
Non-overlapping sets~$E_{i}^{s}$ & Orthogonal projector~$E_{i}^{s}$\tabularnewline\addlinespace
\midrule 
\addlinespace
 & \tabularnewline\addlinespace
\midrule 
\addlinespace
 & \tabularnewline\addlinespace
\midrule 
\addlinespace
 & \tabularnewline\addlinespace
\bottomrule
\end{tabular}
\end{table}
\footnote{\yutsung{functor? Galois connection?}},

we would prove some lemmas which can help us move the classical results
to quantum.
